---
title: "SLF PA and MD Acreage Data 2019"
author: "Chad J. Bronson"
date: "10/23/2019"
output:
  rmdformats::readthedown:
    highlight: tango
---
```{r, echo=FALSE, message=FALSE}
# calling libraries and reading in files (all excel in this instance)
library(ggplot2)
library(prettydoc)
library(formattable)
library(plotly)
library(scales)
library(readxl)
paacres <- read_excel("C:/Users/cbron/OneDrive/Documents/Work/paacres.xlsx", sheet = "ep")
library(readxl)
ct <- read_excel("C:/Users/cbron/OneDrive/Documents/Work/ct.xlsx")
library(readxl)
property_type_table <- read_excel("C:/Users/cbron/OneDrive/Documents/Work/property type table.xlsx")
library(readxl)
patoacres <- read_excel("C:/Users/cbron/OneDrive/Documents/Work/patoacres.xlsx")
library(readxl)
pamdacrescomparison <- read_excel("C:/Users/cbron/OneDrive/Documents/Work/pamdacrescomparison.xlsx")
```
## Tables

### Acreage totals by state, 2019:
```{r, echo=FALSE}
# formattable() is a package for creating better tables than the default table()
formattable(pamdacrescomparison, align = "l")
```

### Property type / use information, PA only:
```{r, echo=FALSE}
# formattable() table cration
formattable(property_type_table, align = "l", list("Acres" = color_tile("white", "pink"),
                                                   "Saplings" = color_tile("white", "pink"),
                                                   "Herbicide Treatments" = color_tile("white", "pink"),
                                                   "Insecticide Treatments" = color_tile("white", "pink"),
                                                   "Canopy Cover" = color_tile("white", "pink"),
                                                   "Completion Time (hours)" = color_tile("white", "pink"),
                                                   "Number of Parcels" = color_tile("white", "pink"),
                                                   "Price ($)" = color_tile("white", "pink")))
```

### Task Order information, PA only:
Key: 

E# = Easton task order number,
G# = Glenside task order number,
H# = Harrisburg task order number.

```{r, echo=FALSE}
# Another formattable table
formattable(patoacres, align = "l", list("Acres" = color_tile("white", "light green"),
                                         "Saplings" = color_tile("white", "light green"),
                                         "Herbicide Treatments" = color_tile("white", "light green"),
                                         "Insecticide Treatments" = color_tile("white", "light green"),
                                         "Canopy Cover (acres)" = color_tile("white", "light green"),
                                         "Completion Time (hours)" = color_tile("white", "light green"),
                                         "Number of Parcels" = color_tile("white", "light green"),
                                         "Price ($)" = color_tile("white", "light green")))
```


### Acreage vs Task Order
```{r, echo=FALSE}
# bar chart using plotly...ggplotly(ggplot()) initializes plotly to make the visuals interactive.
ggplotly(ggplot(ct) + aes(x = ct$`Task Order`,
                          y = ct$Acres, fill = ct$`Task Order`) +
           geom_bar(stat = "identity") +
           xlab("Task Order, PA") +
           ylab("Acres") +
           ggtitle("Parcel Acreage per Task Order, PA") +
           theme(plot.title = element_text(hjust = 0.5)) +
           theme(axis.text.x = element_text(size = 7)) +
           labs(fill = "Task Order") +
           theme(axis.text.x = element_text(angle = 90)) + 
           scale_y_continuous(breaks = seq(0, 3000, 500)))
```

### Price vs Task Order
```{r, echo=FALSE}
# bar chart. Indicated by geom_bar()
ggplotly(ggplot(ct) + aes(x = ct$`Task Order`,
                          y = ct$`Price ($)`, fill = ct$`Task Order`) +
           geom_bar(stat = "identity") +
           xlab("Task Order, PA") + ylab("Price") +
           ggtitle("Price by Task Order, PA") +
           theme(plot.title = element_text(hjust = 0.5))+
           theme(axis.text.x = element_text(size = 7)) +
           scale_y_continuous(breaks = seq(0, 200000, 25000), labels = dollar_format(accuracy = NULL, scale = 1)) +
           labs(fill = "Task Order") +
           theme(axis.text.x = element_text(angle = 90)))
```

### Acreage by Task Order with completion time:

The lighter the shade of blue the more time it took to complete the task order.
```{r, echo=FALSE}
ggplotly(ggplot(ct) + aes(x = ct$`Task Order`,
                          y = ct$Acres, fill = ct$`Completion Time (hours)`) +
           geom_bar(stat = "identity") + xlab("Task Order, PA") + ylab("Acres") +
           ggtitle("Acreage and Price by Task Order, PA") + theme(plot.title = element_text(hjust = 0.5)) +
           theme(axis.text.x = element_text(size = 7)) + labs(fill = "Completion Time (hr)") +
           theme(axis.text.x = element_text(angle = 90)) + scale_y_continuous(breaks = seq(0, 3000, 500)))
```

## Simple Linear Regression Models:
Note: Regression models are meant to be helpful tools for prediction of future values based on all past data in the data set. The line (and equation) are not meant to match the actual x,y point on the graph. The line is a mirror of the variance of all data points. In other words, the line is chosen based on the smallest average difference between all points and the line. We'll be able to see which variables seem to effect completion time the most. If you hover over the line with your mouse you can see how the x variable(s) effect completion time. These can be done by task order, or on an individual parcel basis. Both will yield similar results.

### Task order calculations examples:

### Completion time per task order vs Parcel acreage per task order:
Estimated completion time = 13.42hr + 0.05935x

R$^2$= 0.53. This value measures how close the data is to the line. The closer to 1 the better the fit of the line, the closer to 0 the less correlation. If its close to 1 it will be a better predictor than if it close to 0.

Example. Easton task order 1: y = 13.42 + 0.05935x:

estimated completion time(y) = 13.42 + 0.05935(20.08)

estimated completion time = 14.60hrs. The actual completion time was 14.68 hours.

If we look at task order 5 for easton it took 106.32 hours to complete. The estimated time to complete task order 5 was 28.97 hours. This shows how much variability there is in acres vs completion time. 
```{r, echo=FALSE}
# This is a linear regression. Two geom_ calls are made here to plot the points and then the line to fit.
# Doing a linear/multiple linear regression without a visual would look like:
#            variable name <- lm(y_axis_variable ~ x1_axis_variable, x2_axis_variable, etc...)
ggplotly(ggplot(ct) + aes(x = ct$Acres,
                          y = ct$`Completion Time (hours)`) +
           geom_point(stat = "identity") + geom_smooth(method = "lm", se = FALSE) +
           xlab("Acres per Task Order") + ylab("Completion Time (hours)") +
           ggtitle("Task Order Acreage and Completion Time (hours)") +
           scale_x_continuous(breaks = seq(0, 3000, 500) + theme(plot.title = element_text(hjust = 0.5))))
```

### Completion Time vs Canopy Cover per task order:

Completion Time = 25.95 + 0.05599x

R$^2$= 0.06

Based on the 2019 PA data, this is a poor predictor with very weak correlation.
```{r, echo=FALSE}
ggplotly(ggplot(ct) + aes(x = ct$`Canopy Cover`,
                          y = ct$`Completion Time (hours)`) +
           geom_point(stat = "identity") +
           geom_smooth(method = "lm", se = FALSE) +
           xlab("Canopy Cover per Task Order PA") +
           ylab("Completion Time (hours)") +
           ggtitle("Canopy Cover per Task Order and Completion Time (hours)") +
           scale_x_continuous(breaks = seq(0, 1200, 200)) +
           theme(plot.title = element_text(hjust = 0.5)))
```

### Completion Time vs Price (task order):
estimated completion time = -2.3 + 0.000978(price)

R$^2$ = 0.72
```{r, echo=FALSE}
ggplotly(ggplot(ct) + aes(x = ct$`Price ($)`,
                          y = ct$`Completion Time (hours)`) +
           geom_point(stat = "identity") + geom_smooth(method = "lm", se = FALSE) +
           xlab("Price per Task Order, PA") + ylab("Completion Time (hours)") +
           ggtitle("Price per Task Order and Completion Time (hours)") +
           theme(plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(breaks = seq(0, 200000, 20000),
                              labels = dollar_format(accuracy = NULL, scale = 1)))
```

### Individual parcel calculations:

### Saplings Treated vs Completion Time for **every parcel** (the following will show equations on a per parcel basis, not per task order):

completion time = 1.98 + 0.027x

R$^2$ = 0.37 
```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$`Saplings Treated`,
                               y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") +
           geom_smooth(method = "lm", se = FALSE) +
           xlab("Saplings Treated per Parcel, PA") +
           ylab("Completion Time (hours)") +
           ggtitle("Saplings Treated per Parcel and Completion Time (hours)") +
           scale_x_continuous(breaks = seq(0, 4000, 500)) +
           theme(plot.title = element_text(hjust = 0.5)))
```

### Completion Time vs Acres for every parcel:

Completion Time = 1.5 + 0.063x

R$^2$= 0.56
```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$Acres,
                               y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") +
           geom_smooth(method = "lm", se = FALSE) +
           xlab("Acres Treated per Parcel, PA") +
           ylab("Completion Time (hours)") +
           ggtitle("Acres Treated per Parcel and Completion Time (hours)") +
           theme(plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(breaks = seq(0, 3000, 500)))
```

### Completion Time vs Canopy Cover:
estimated completion time = 2.98 + 0.07x

R$^2$ = 0.07
```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$`Canopy Cover (acres)`,
                               y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") +
           geom_smooth(method = "lm", se = FALSE) +
           xlab("Canopy Cover per Parcel, PA") +
           ylab("Completion Time (hours)") +
           ggtitle("Canopy Cover per Parcel and Completion Time (hours)") +
           theme(plot.title = element_text(hjust = 0.5)))
```

### Completion Time vs Herbicide Treatments (at least 1in DBH) for every parcel:

estimated completion time = 0.53 + 0.052x

R$^2$ = 0.71

```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$`Herbicide Applications 1-5.99in in DBH`
                               , y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") + geom_smooth(method = "lm", se = FALSE) +
           xlab("Herbicide Applications, PA") + ylab("Completion Time (hours)") +
           ggtitle("Herbicide Applications (no saplings) and Completion Time") +
           scale_x_continuous(breaks = seq(0, 3000, 500)) +
           theme(plot.title = element_text(hjust = 0.5)))
```

### Completion Time vs Insecticide Applications for every property

estimated completion time = 1.69 + 0.184x

R$^2$ = 0.164

```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$`Insecticide Applications 6+ in DBH`,
                               y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") + geom_smooth(method = "lm", se = FALSE) +
           xlab("Insecticide Applications, PA") + ylab("Completion Time (hours)") +
           ggtitle("Insecticide Applications and Completion Time") +
           theme(plot.title = element_text(hjust = 0.5)))
```

### Completion time vs Price:
estimated completion time = -0.4 + 0.00096x

R$^2$ = 0.7983

```{r, echo=FALSE}
ggplotly(ggplot(paacres) + aes(x = paacres$`Price ($)`,
                               y = paacres$`Completion Time (hours)`) +
           geom_point(stat = "identity") +
           geom_smooth(method = "lm", se = FALSE) +
           xlab("Price") + ylab("Completion Time (hours)") +
           ggtitle("Completion Time (hours) and Price ($)") +
           scale_x_continuous(breaks = seq(0, 200000, 20000),
                              labels = dollar_format(accuracy = NULL, scale = 1)) +
           theme(plot.title = element_text(hjust = 0.5)))
```

If we examine all of the above plots we would see that Price has the highest R$^2$ value (0.7983) and would give us the best estimate of how long it would take to complete a property if we are to only use 1 predictor variable (not by task order, but by each property).

## Multiple Linear Regression Equation:

When we take all of these variables vs time and include them in the same equation we get the following:

**estimated completion time = -0.0134 + 0.0364(acres) - 0.0333(canopy cover) - 0.00458(saplings) - 0.000551(herbicide applications of at least 1in DBH) - 0.0833(insecticide applications) + 0.0009403(price)**

Adjusted R$^2$ = 0.93

This will give the closest and most consistent estimate overall. However since the data set is still small it will only be a ballpark estimate. These models are only applicable based on 2019 PA data, and as more parcels are added and treated this would need to be updated every time. The more data, the more accurate this can be if there truly is any correlation with the predictor variables and completion time.
